---
title: <center><font size="6"><b>Detecting 55 kHz USVs with bedding</b></font></center>
subtitle: <center><font size="4"><b>Rat ultrasonic vocalizations</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 3
    df_print: kable
    toc_float:
      collapsed: yes
      smooth_scroll: yes
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

```{r packages, message = FALSE, warning = FALSE, echo = FALSE, eval = TRUE, include = FALSE}

## add 'developer/' to packages to be installed from github
x <- c("devtools", "maRce10/warbleR", "bioacoustics", "pbapply", "Rraven", "parallel", "viridis", "rfigshare", "svMisc", "readxl", "ranger", "kableExtra", "DT", "maRce10/ohun", "ggplot2")

aa <- lapply(x, function(y) {
  
  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]
  
  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  devtools::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

```

```{r functions and global parameters, eval = TRUE, echo = FALSE}

warbleR_options(
  wav.path = "~/Dropbox/Recordings/ratas_cin/recs/200_kHz_bedding/"
)


opts_knit$set(root.dir = "..")

source("~/Dropbox/R_package_testing/ohun/R/split_acoustic_data.R")
source("~/Dropbox/R_package_testing/warbleR/warbleR/R/overlapping_sels.R")
source("~/Dropbox/R_package_testing/warbleR/warbleR/R/pblapply_wrblr_int.R")
source("~/Dropbox/R_package_testing/ohun/R/label_detection.R")
source("~/Dropbox/R_package_testing/ohun/R/filter_detection.R")
source("~/Dropbox/R_package_testing/warbleR/warbleR/R/overlapping_sels.R")

merged_overlaps <- function(X){
  
  # merged overlapping selections
  ov_sls <- overlapping_sels(X)
  
  ov_sls <- do.call(rbind, lapply(unique(ov_sls$ovlp.sels), function(x){
         
    if (!is.na(x)){
       
       Y <- ov_sls[ov_sls$ovlp.sels == x & !is.na(ov_sls$ovlp.sels), ]
       Y$end[1] <- max(Y$end)
       Y$bottom.freq[1] <- min(Y$bottom.freq)
       Y$top.freq[1] <- max(Y$top.freq)
       Y <- Y[1, ]
       } else
        Y <- ov_sls[is.na(ov_sls$ovlp.sels), ]
      return(Y)
  }))
  
  ov_sls$ovlp.sels <- NULL
   
  return(ov_sls) 
 
}

```

# Optimize detection for 55 kHz single cut with more signals

```{r optimize detection for 55 kHz single 5 min cut, eval = FALSE}

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv", stringsAsFactors = FALSE)

sels_55 <- split_sels[split_sels$peak.freq > 30, ]

table(sels_55$sound.files)

sels_55 <- split_sels[split_sels$sound.files == "TH-27.wav", ]

# merged overlapping selections
sels_55 <- merged_overlaps(sels_55)

oed <- optimize_energy_detector(reference = sels_55, threshold = seq(0.01, 0.03, 0.005), min.duration = c(0.005, 0.01, 0.001), ssmooth = c(1, 2.5, 5, 10), hold.time = c(0.0025, 0.005, 0.01, 0.05, seq(0.08, 0.1, 0.01)), path = .Options$warbleR$path, thinning = c(0.5), parallel = 4, bp = c(35, 90), max.duration = c(0.15, 0.2, 0.22), previous.output = oed)

saveRDS(oed, "./data/processed/optimization_55kHz_with_bedding.RDS")

```

```{r results optimize detection for single 5 min cut, eval = TRUE}

optim_ad <- readRDS("./data/processed/optimization_55kHz_with_bedding.RDS")

# print dynamic table
oa_DT <- datatable(optim_ad, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(optim_ad, is.numeric), 3)

```

# Optimize detection for 55 kHz 5 cuts with more signals

```{r optimize detection for 55 kHz 5 cuts, eval = FALSE}

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv")

sels_55 <- split_sels[split_sels$peak.freq > 30, ]

# merged overlapping selections
sels_55 <- merged_overlaps(sels_55)

tab <- table(sels_55$sound.files)
sels_55 <- sels_55[sels_55$sound.files %in% names(tab)[tab > 240], ]


# best tuning parameters 
oed5 <- optimize_energy_detector(reference = sels_55, files =  unique(sels_zoo55$sound.files), threshold = c(0.025, 0.03), min.duration = c(0.001), ssmooth = c(1, 2.5, 5), hold.time = c(0.005, 0.003), path = .Options$warbleR$path, thinning = 0.5, parallel = 1, bp = c(35, 90), max.duration = c(0.15, 0.2, 0.22), by.sound.file = TRUE)

saveRDS(oed5, "./data/processed/optimization_55kHz_5_cuts_with_bedding.RDS")

```

```{r results optimize detection for 55 kHz 5 cuts, eval = TRUE}

optim_ad <- readRDS("./data/processed/optimization_55kHz_5_cuts_with_bedding.RDS")

optim_ad <- summarize_diagnostic(optim_ad, time.diagnostics = TRUE)

oa_DT <- datatable(optim_ad, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(optim_ad, is.numeric), 3)
 
```

# Optimal detection diagnostic 55 kHz 5 cuts

```{r optimize detection for 55 kHz 5 cuts try selected tuning parameters, eval = FALSE}

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv")

sels_55 <- split_sels[split_sels$peak.freq > 30, ]

# merged overlapping selections
sels_55 <- merged_overlaps(sels_55)

tab <- table(sels_55$sound.files)
sels_55 <- sels_55[sels_55$sound.files %in% names(tab)[tab > 240], ]


# best tuning parameters 
ed5 <- energy_detector(files =  unique(sels_55$sound.files), threshold = 0.025, min.duration = 0.001, ssmooth = 1, hold.time = 0.005, path = .Options$warbleR$path, thinning = 0.5, parallel = 1, bp = c(35, 90), max.duration = 0.22)

# fed4
saveRDS(ed5, "./data/processed/optimal_detection_55kHz_5_cuts_with_bedding.RDS")

```

```{r diagnose detection for 55 kHz 5 cuts try selected tuning parameters results by sound file, eval = FALSE}
ed5 <- readRDS("./data/processed/optimal_detection_55kHz_5_cuts_with_bedding.RDS")

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv")

sels_55 <- split_sels[split_sels$peak.freq > 30, ]

# merged overlapping selections
sels_55 <- merged_overlaps(sels_55)

tab <- table(sels_55$sound.files)
sels_55 <- sels_55[sels_55$sound.files %in% names(tab)[tab > 240], ]
led5 <- label_detection(reference = sels_55, detection = ed5, parallel = 10, pb = FALSE)

led5 <- filter_detection(led5, parallel = 10, pb = FALSE)

optim_ad_bs <- diagnose_detection(reference = sels_55, detection = led5, by.sound.file = TRUE)

optim_ad <- diagnose_detection(reference = sels_55, detection = led5, by.sound.file = FALSE)

saveRDS(list(call_5 = attributes(ed5)$call, filtered_detect_5 = led5, optim_ad_bs_5 = optim_ad_bs, optim_ad_5 = optim_ad), "./data/processed/filtered_detection_55Khz_5_cuts_with_bedding.RDS")

```

## By sound file
```{r optimize detection for 55 kHz 5 cuts print results by sound files, eval = TRUE, warning=FALSE}

attach(readRDS("./data/processed/filtered_detection_55Khz_5_cuts_with_bedding.RDS"))

call_5

oa_DT <- datatable(optim_ad_bs_5, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(optim_ad_bs_5, is.numeric), 3)

```

## Summarized

```{r optimize detection for 55 kHz 5 cuts print results, eval = TRUE, warning=FALSE}

oa_DT <- datatable(optim_ad_5, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(optim_ad_5, is.numeric), 3)

  
```

# Detection over all files

# Optimize detection

```{r optimize detection for 55 kHz all cuts, eval = FALSE}

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv")

sels_55 <- split_sels[split_sels$peak.freq > 30, ]

# merged overlapping selections
sels_55 <- merged_overlaps(sels_55)

# best tuning parameters 
oed_all <- optimize_energy_detector(reference = sels_55, files =  unique(sels_55$sound.files), threshold = c(0.025, 0.03), min.duration = c(0.001), ssmooth = c(1, 2.5, 5), hold.time = c(0.005, 0.003), path = .Options$warbleR$path, thinning = 0.5, parallel = 1, bp = c(35, 90), max.duration = c(0.15, 0.2, 0.22), by.sound.file = FALSE)


saveRDS(oed_all, "./data/processed/optimization_55kHz_all_cuts_with_bedding.RDS")

```

```{r optimize detection for 55 kHz all files cuts print results , eval = TRUE, warning=FALSE}

oed_all <- readRDS("./data/processed/optimization_55kHz_all_cuts_with_bedding.RDS")

oa_DT <- datatable(oed_all, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(oed_all, is.numeric), 3)

```

```{r Run detection over all files, eval=FALSE}

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv")

# merged overlapping selections
split_sels <- merged_overlaps(split_sels)

# best tuning parameters 
ed_all <-  energy_detector(files = unique(split_sels$sound.files), threshold = 0.025, min.duration = 0.001, ssmooth = 1, hold.time = 0.003, path = .Options$warbleR$path, thinning = 0.5, parallel = 1, bp = c(35, 90), max.duration = 0.2)

saveRDS(ed_all, "./data/processed/optimal_detection_55kHz_all_cuts_with_bedding.RDS")

```

```{r diagnose detection all 55 kHz cuts try selected tuning parameters results by sound file, eval = FALSE}
ed_all <- readRDS("./data/processed/optimal_detection_55kHz_all_cuts_with_bedding.RDS")

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv")

sels_55 <- split_sels[split_sels$peak.freq > 30, ]

# merged overlapping selections
sels_55 <- merged_overlaps(sels_55)

label_ed_all <- label_detection(reference = sels_55, detection = ed_all, parallel = 10, pb = FALSE)
filter_ed_all <- filter_detection(label_ed_all, parallel = 10, pb = FALSE)

optim_ad_bs_all <- diagnose_detection(reference = sels_55, detection = filter_ed_all, by.sound.file = TRUE, pb = FALSE)

optim_ad_all <- diagnose_detection(reference = sels_55, detection = ed_all, pb = FALSE)

saveRDS(list(call = attributes(ed_all)$call, filter_ed_all = filter_ed_all, optim_ad_bs_all = optim_ad_bs_all, optim_ad_all = optim_ad_all), "./data/processed/filtered_detection_55Khz_all_cuts_with_bedding.RDS")

```

## By sound file

```{r print diagnose detection all 55 kHz cuts results by sound file, eval = TRUE, warning=FALSE}

attach(readRDS("./data/processed/filtered_detection_55Khz_all_cuts_with_bedding.RDS"))

call

oa_DT <- datatable(optim_ad_bs_all, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(optim_ad_bs_all, is.numeric), 3)

```

## Summarized

```{r print diagnose detection all 55 kHz cuts results, eval = TRUE}

oa_DT <- datatable(optim_ad_all, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(optim_ad_all, is.numeric), 3)

  
```

# Random forest classification
```{r measure acoustic parameters and run random forest, eval = FALSE}

# measure spectrographic parameters
spectral_parameters <- spectro_analysis(filter_ed_all, bp = c(35, 85), fast = TRUE, ovlp = 70, parallel = 5)


spectral_parameters$class <- filter_ed_all$detection.class

# spectral_parameters <- data.frame(spectral_parameters, mfccs[, !names(spectral_parameters) %in% c("sound.files", "selec")])

spectral_parameters$class[spectral_parameters$class != "false.positive"] <- "true.positive"

# make it a factor for ranger to work 
spectral_parameters$class <- as.factor(spectral_parameters$class)
  


  # run RF model spectral and cepstral parameters
  rfm <-
    ranger(
      class ~ .,
      data = spectral_parameters[, !names(spectral_parameters) %in% c("sound.files", "selec")],
      num.trees = 10000,
      importance = "impurity",
      seed = 10
    )

  # table(lab_detec$detection.class)
filter_ed_all$pred.class <- rfm$predictions  
  

saveRDS(list(rfm = rfm, filter_ed_all_rf = filter_ed_all), "./data/processed/data_and_model_random_forest_55kHz_cuts_with_bedding.RDS")  

```

```{r random forest results, eval = TRUE}
attach(readRDS("./data/processed/data_and_model_random_forest_55kHz_cuts_with_bedding.RDS"))

rfm
```

# Final results
Diagnostic after random forest classification:
```{r random forest diagnostic, eval = TRUE}

positive_detec <- filter_ed_all_rf[filter_ed_all_rf$pred.class == "true.positive", ]

temp_detec <- positive_detec
temp_detec$detection.class <- "true.positive"

split_sels <- read.csv("./data/processed/split_manual_annotations_5min_bedding.csv")

sels_55 <- split_sels[split_sels$peak.freq > 30, ]

# merged overlapping selections
sels_55 <- merged_overlaps(sels_55)

diag <- diagnose_detection(reference = sels_55, detection = temp_detec, pb = FALSE)

# print dynamic table
oa_DT <- datatable(diag, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

formatRound(table = oa_DT, columns = sapply(diag, is.numeric), 3)
```

 Black line = 1:1
 gray line = model slope
```{r random forest results plots, eval = TRUE}

obs_count <- tapply(sels_55$sound.files, sels_55$sound.files, length)
pred_count <- tapply(positive_detec$sound.files, positive_detec$sound.files, length)
sound_files <- unique(split_sels$sound.files)


# add those missing in predicted
names_pred_0 <- setdiff(c(names(obs_count),sound_files), names(pred_count))
pred_0 <- rep(0, length(names_pred_0))
names(pred_0) <- names_pred_0
pred_count <- c(pred_count, pred_0)

# add those missing in observed
names_obs_0 <- setdiff(c(names(pred_count), sound_files), names(obs_count))
obs_0 <- rep(0, length(names_obs_0))
names(obs_0) <- names_obs_0
obs_count <- c(obs_count, obs_0)

# order by name
pred_count <- pred_count[order(names(pred_count))]
obs_count <- obs_count[order(names(obs_count))]

# put both in a single data frame
df <- data.frame(sound.files = names(obs_count), observed = obs_count, predicted = pred_count)

# plot
ggplot(df, aes(x = observed, y = predicted)) +
  geom_point(color = viridis(10, alpha = 0.4)[2], size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 150, y = 550, label = paste("r =", round(cor(obs_count, pred_count), 3)), size = 8) + 
  geom_smooth(method = "lm", se = FALSE, col = "gray") +
  theme_classic(base_size = 20)


# print best fit lm model
(lm(pred_count ~ obs_count))

```

```{r splitting in 1 minute clips, eval = FALSE}
# split
split_obs <- split_acoustic_data(path = .Options$warbleR$path, X = sels_55, only.sels = TRUE, sgmt.dur = 60, pb = FALSE, parallel = 10)

split_pred <- split_acoustic_data(path = .Options$warbleR$path, X = as.data.frame(positive_detec), only.sels = TRUE, sgmt.dur = 60, pb = FALSE, parallel = 10)

obs_count <- tapply(split_obs$sound.files, split_obs$sound.files, length)
pred_count <- tapply(split_pred$sound.files, split_pred$sound.files, length)
sound_files <- unique(split_sels$sound.files)


# add those missing in predicted
names_pred_0 <- setdiff(c(names(obs_count),sound_files), names(pred_count))
pred_0 <- rep(0, length(names_pred_0))
names(pred_0) <- names_pred_0
pred_count <- c(pred_count, pred_0)

# add those missing in observed
names_obs_0 <- setdiff(c(names(pred_count), sound_files), names(obs_count))
obs_0 <- rep(0, length(names_obs_0))
names(obs_0) <- names_obs_0
obs_count <- c(obs_count, obs_0)

# order by name
pred_count <- pred_count[order(names(pred_count))]
obs_count <- obs_count[order(names(obs_count))]

saveRDS(list(pred_count = pred_count, obs_count = obs_count), "observed_and_predicted_1_min_clips.RDS")
```


```{r, eval = FALSE}
attach(readRDS("./data/processed/observed_and_predicted_1_min_clips.RDS"))

# put both in a single data frame
df <- data.frame(sound.files = names(obs_count), observed = obs_count, predicted = pred_count)

df$sum <- df$observed + df$predicted

# plot
ggplot(df[df$sum > 0, ], aes(x = observed, y = predicted)) +
  geom_point(color = viridis(10, alpha = 0.4)[2], size = 3) +
  geom_abline(slope = 1, intercept = 0) +
  annotate("text", x = 20, y = 50, label = paste("r =", round(cor(obs_count, pred_count), 3)), size = 8) + 
  geom_smooth(method = "lm", se = FALSE, col = "gray") +
  theme_classic(base_size = 20)


# print best fit lm model
(lm(pred_count ~ obs_count))


```

---

<font size="4">Session information</font>

```{r session info, echo=F, eval = TRUE}

sessionInfo()

```
