---
title: <center><font size="6"><b>Format USVs recordings</b></font></center>
subtitle: <center><font size="4"><b>Rat ultrasonic vocalizations</b></font></center>
author: <center><font size="4"><a href="http://marceloarayasalas.weebly.com/">Marcelo Araya-Salas, PhD</a></font></center>
date: <center>`r format(Sys.Date(), "%d-%m-%Y")`</center>
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 3
    df_print: kable
    toc_float:
      collapsed: yes
      smooth_scroll: yes
fontsize: 12pt 
editor_options: 
  chunk_output_type: console
---

# Load packages

```{r packages, message = FALSE, warning = FALSE, echo = FALSE, eval = TRUE, include = FALSE}

## add 'developer/' to packages to be installed from github
x <- c("devtools", "maRce10/warbleR", "Rraven", "DT")

aa <- lapply(x, function(y) {
  
  # get pakage name
  pkg <- strsplit(y, "/")[[1]]
  pkg <- pkg[length(pkg)]
  
  # check if installed, if not then install 
  if (!pkg %in% installed.packages()[,"Package"])  {

      if (grepl("/", y))  devtools::install_github(y, force = TRUE) else
    install.packages(y) 
    }

  # load package
  try(require(pkg, character.only = T), silent = T)
})

```

```{r functions and global parameters, eval = TRUE, echo = FALSE}

opts_knit$set(root.dir = "..")

source("~/Dropbox/R_package_testing/ohun/R/split_sound_files.R")

```

# Formatting steps

- Make sure names in files and annotations match
- Downsample to 200 kHz
- Split sound files in 5 minute clips (only for no-bedding files)
- Split annotations to refer to 5 minute clips (only for no-bedding files)

# Format no-bedding recording and annotations

```{r read manual selections no bedding, eval = FALSE}

mnl_sls <- imp_raven(path = "./data/raw/no_bedding_annotations", warbler.format = TRUE, all.data = TRUE, name.from.file = TRUE, ext.case = "upper", unread = TRUE)

unread_l <- lapply(file.path("./data/raw/no_bedding_annotations/", mnl_sls$unread_files), function(x) {
  X <- read.table(x, sep = ",", header = TRUE)
  X$selec.file <- X$sound.files <- basename(x)
  return(X)
  })

# remove 4 and 5 as those have only 1 selection and no frequency info
unread_l <- unread_l[1:3]

unread <- do.call(rbind, unread_l)
unread$type <- unread$subtype <- NA

unread <- relabel_colms(unread, hz.to.khz = TRUE)
unread <- sort_colms(unread)

head(unread)

mnl_sls <- sort_colms(mnl_sls$selections)

names(mnl_sls)
names(unread)
mnl_sls$`Begin File` <- NULL

mnl_sls <- rbind(mnl_sls, unread)

mnl_sls$sound.files <- gsub("A_AMPH_15-45_estres_parte2_AUDIO |A_AMPH_15-45_estres_parte2_Audio |A_AMPH_15-45_estres_parte2_AUDIO_", "", mnl_sls$sound.files)

mnl_sls$sound.files <- gsub("T000003.WAV", "T0000003.WAV", mnl_sls$sound.files)
mnl_sls$sound.files <- gsub("T000009.WAV", "T0000009.WAV", mnl_sls$sound.files)
mnl_sls$sound.files <- gsub("T000011.WAV", "T0000011.WAV", mnl_sls$sound.files)
mnl_sls$sound.files <- gsub("T000012.WAV", "T0000012.WAV", mnl_sls$sound.files)
mnl_sls$sound.files <- gsub("T000013.WAV", "T0000013.WAV", mnl_sls$sound.files)
mnl_sls$sound.files <- gsub("T000014.WAV", "T0000014.WAV", mnl_sls$sound.files)

mnl_sls$sound.files <- gsub(".txt$", ".WAV", mnl_sls$sound.files)

unique(mnl_sls$sound.files)

mnl_sls$selec <- 1:nrow(mnl_sls)

cs <- check_sels(mnl_sls)

# unique(cs$sound.files[grep("sound file not found", cs$check.res)])

# table(cs$check.res)

mnl_sls <- mnl_sls[cs$check.res == "OK", ]

cs <- check_sels(mnl_sls)


mnl_sls$bottom.freq[mnl_sls$selec == 2474] <- 48.9885
mnl_sls$top.freq[mnl_sls$selec == 2474] <- 73.8728
mnl_sls$bottom.freq[mnl_sls$selec == 2198] <- 59.6217
mnl_sls$top.freq[mnl_sls$selec == 2198] <- 74311.4
mnl_sls$top.freq[mnl_sls$selec == 1852] <- 111473.9

mnl_sls$top.freq[mnl_sls$top.freq > 100] <- 98

mnl_sls <- mnl_sls[mnl_sls$bottom.freq > 20,]

# export raven file to double check
# exp_raven(mnl_sls, path = "./data/processed", file.name = "pooled_manual_annotations", sound.file.path = .Options$warbleR$path)


```

```{r, import annotations fix and split sound files no bedding, eval = FALSE}

# fix_path(path = "./data/processed/", sound.file.col = "Begin File", new.begin.path = "/home/m/Dropbox/Recordings/ratas_cin/full_recs/converted_sound_files")

manual_annot <- imp_raven("./data/processed/", warbler.format = TRUE, files = "fixed_pooled_manual_annotations_JE.txt")

cs <- check_sels(manual_annot, path = "~/Dropbox/Recordings/ratas_cin/recs/original_with_bedding/")

wav_info()

# fix_wavs(samp.rate = 200, bit.depth = 16)

# split in 5 min segments
split_sels <- split_sound_files(sgmt.dur = 5 * 60, only.sels = FALSE, X = manual_annot, parallel = 4)

cs <- check_sels(manual_annot, path = "~/Dropbox/Recordings/ratas_cin/recs/5-min_clips_no_bedding/")

split_sels <- split_sels[cs$check.res == "OK", ]

sp <- spectro_analysis(split_sels, wl = 512, parallel = 10, path = "~/Dropbox/Recordings/ratas_cin/recs/5-min_clips_no_bedding/")

split_sels$peak.freq <- sp$meanpeakf
split_sels$duration <- split_sels$end - split_sels$start


split_sels <- sig2noise(split_sels, mar = 0.005, parallel = 10, path = "~/Dropbox/Recordings/ratas_cin/recs/5-min_clips_no_bedding/")

write.csv(split_sels, "./data/processed/split_manual_annotations_5min_no_bedding.csv", row.names = FALSE)

# export raven file to double check
exp_raven(split_sels, path = "./data/processed", file.name = "split_manual_annotations_5min", sound.file.path = .Options$warbleR$path)

```

## 5 min clips metadata no-bedding
```{r}

wi <- wav_info(path = "~/Dropbox/Recordings/ratas_cin/recs/5-min_clips_no_bedding/")

datatable(wi, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

```

---

# Format bedding recording and annotations

```{r read and format manual selection and recordings bedding, eval = FALSE}

mnl_sls <- imp_raven(path = "./data/raw/with_bedding", warbler.format = TRUE, all.data = TRUE, name.from.file = TRUE, ext.case = "upper", unread = FALSE)

mnl_sls$sound.files <- gsub("\\.WAV", ".wav", mnl_sls$sound.files)


fls <- list.files("~/Dropbox/Recordings/ratas_cin/recs/original_with_bedding/")

file.rename(from = file.path("~/Dropbox/Recordings/ratas_cin/recs/original_with_bedding/", fls), to = file.path("~/Dropbox/Recordings/ratas_cin/recs/original_with_bedding/", paste0(substr(fls, 0, 5), ".wav")))

mnl_sls$sound.files <- paste0(substr(mnl_sls$sound.files, 0, 5), ".wav")

cs <- check_sels(mnl_sls, path = "~/Dropbox/Recordings/ratas_cin/recs/original_with_bedding/")

table(cs$check.res)

mnl_sls <- mnl_sls[cs$check.res == "OK", ]

mnl_sls$top.freq[mnl_sls$top.freq > 100] <- 98

mnl_sls <- mnl_sls[mnl_sls$bottom.freq > 20,]

# export raven file to double check
exp_raven(mnl_sls, path = "./data/processed", file.name = "pooled_manual_annotations_with_bedding", sound.file.path = "~/Dropbox/Recordings/ratas_cin/recs/original_with_bedding/")

```

```{r, import annotations fix and split sound files no bedding, eval = FALSE}

# fix_path(path = "./data/processed/", sound.file.col = "Begin File", new.begin.path = "/home/m/Dropbox/Recordings/ratas_cin/full_recs/converted_sound_files")

manual_annot <- imp_raven(path = "./data/processed", files = "pooled_manual_annotations_with_bedding.txt", warbler.format = TRUE, all.data = FALSE)


cs <- check_sels(manual_annot, path = "~/Dropbox/Recordings/ratas_cin/recs/original_with_bedding/")

wav_info()

# fix_wavs(samp.rate = 200, bit.depth = 16)

sp <- spectro_analysis(manual_annot, wl = 512, parallel = 10, path = "~/Dropbox/Recordings/ratas_cin/recs/200_kHz_bedding/")

manual_annot$peak.freq <- sp$meanpeakf
manual_annot$duration <- manual_annot$end - manual_annot$start

manual_annot <- sig2noise(manual_annot, mar = 0.005, parallel = 10, path = "~/Dropbox/Recordings/ratas_cin/recs/200_kHz_bedding/")

write.csv(manual_annot, "./data/processed/split_manual_annotations_5min_bedding.csv", row.names = FALSE)

```


## Sound files metadata no-bedding
```{r}

wi <- wav_info(path = "~/Dropbox/Recordings/ratas_cin/recs/200_kHz_bedding/")

datatable(wi, editable = list(
  target = 'row'
), rownames = FALSE, style = "bootstrap",  filter = 'top', options = list(
  pageLength = 100, autoWidth = TRUE, dom = 'ft'
), autoHideNavigation = TRUE, escape = FALSE)

```


---

<font size="4">Session information</font>

```{r session info, echo=F, eval = TRUE}

sessionInfo()

```
